Analysing the effect of land-use type on the proportion of scouts

### 2021-11-26 analysis/landuse_effect_prop_scouts.Rmd
updated LU map plots

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
# theme_set(
#   theme_classic() +
#     theme(
#       text = element_text(size = rel(42))
#     )
# )

@@

# check the land-use type values sum to 100%
testthat::test_that(
  "land-use type values sum to 100",
  {
    expected <- rep(100, 10)
    ans <- rowSums(rural_lu[, 3:ncol(rural_lu)])
    testthat::expect_identical(expected, ans)
  }
)
```

@@

@@

# make the maps for each of the sites selected in the previous step
site_shape_data <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/agri-rural/",
    ar_marked[i, "site"], "/", ar_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  site_shape_data[[ar_marked[i, "site"]]] <- df
  site_plots[[ar_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df,
      aes(fill = descriptio)
    ) +
    theme_void()
}

@@

loading_plot <- ggplot(argi.rural.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(3)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.25,
    nudge_y = c(0, 0), nudge_x = c(0.5, 0.5), size = rel(3)
  ) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None"
  ) +
  scale_colour_manual(
    values = c(
      "arable" = "red3", "built.up.area" = "orange3",
      "Non.agricultural.unmanaged.green.space" = "seagreen3",
      "water" = "cornflowerblue", other = "grey"
    )
  ) +
  ylim(-0.6, 0.6)

loading_plot

@@

# now the model fit and pls results in a plot of two
mfitpls <- plot_grid(
  ar_fit + theme(plot.margin = unit(c(0., 0., 1., 0.), "cm")),
  loading_plot,
  ncol = 1,
  align = "v",
  labels = c(
    "A", "B"
  )
)

comb
```

@@

@@

# make the maps for each of the sites selected in the previous step
site_shape_data <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/urban/",
    u_marked[i, "site"], "/", u_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  site_shape_data[[u_marked[i, "site"]]] <- df
  site_plots[[u_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df,
      aes(fill = descriptio)
    ) +
    theme_void()
}

@@

loading_plot <- ggplot(urban.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(3)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.25,
    nudge_y = c(-0, -0), nudge_x = c(0.5, 0.5), size = rel(3)
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None"
  ) +
  scale_colour_manual(
    values = c(
      "continuous..central" = "red3", "dense.residential" = "orange3",
      "sparse.residential" = "purple3", "amenity.grassland" = "seagreen3",
      "water" = "cornflowerblue", other = "grey"
    )
  ) +
  scale_y_continuous(breaks = seq(-.9, .9, by = 0.4), limits = c(-1, 1))

loading_plot

@@

comb
```

>>>>

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic()
)

@@

# # check the land-use type values sum to 100%
# testthat::test_that(
#   "land-use type values sum to 100",
#   {
#     expected <- rep(100, 10)
#     ans <- rowSums(rural_lu[, 3:ncol(rural_lu)])
#     testthat::expect_identical(expected, ans)
#   }
# )
```

@@

# load land-use conversion table
luconversions <- read.csv("data/LUCategoryConversion.csv")
agrirural_luconversions <- luconversions %>%
  filter(LU == "agri-rural")

@@

# make the maps for each of the sites selected in the previous step
site_shape_data_ar <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/agri-rural/",
    ar_marked[i, "site"], "/", ar_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  # add major LU categories
  df_merged <- df %>%
    mutate(descriptio = toupper(descriptio)) %>%
    full_join(agrirural_luconversions, by = c("descriptio" = "code"))
  site_shape_data_ar[[ar_marked[i, "site"]]] <- df_merged
  site_plots[[ar_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df_merged,
      aes(fill = category)
    ) +
    theme_void() +
    scale_fill_manual(
      values = c(
        "arable" = "#EE6677", "built-up area" = "#CCBB44",
        "Non-agricultural unmanaged green space" = "#228833",
        "water" = "#4477AA", "pasture" = "#BBBBBB", "woodland" = "#BBBBBB",
        "Non-agricultural managed green space" = "#BBBBBB", "fruit" = "#BBBBBB",
        "OSR" = "#BBBBBB", "other agricultural" = "#BBBBBB"
      )
    )
}

@@

loading_plot <- ggplot(argi.rural.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(3)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.25,
    nudge_y = c(0, 0), nudge_x = c(0.5, 0.5), size = rel(3)
  ) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None"
  ) +
  scale_colour_manual(
    values = c(
      "arable" = "#EE6677", "built.up.area" = "#CCBB44",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  ) +
  ylim(-0.6, 0.6)

@@

# now the model fit and pls results in a plot of two
mfitpls <- plot_grid(
  ar_fit + theme(plot.margin = unit(c(0., 0., 0.2, 0.), "cm")),
  loading_plot,
  ncol = 1,
  align = "v",
  labels = c(
    "A", "B"
  )
)

# all together
comb <- plot_grid(
  mfitpls, sitelumaps,
  nrow = 1, scale = 0.9, align = "h",
  rel_widths = c(1, .5)
)

# MAPS BELLOW PLOT
# first the maps in a column of three
sitelumaps <- plot_grid(
  site_plots[[1]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
  ),
  site_plots[[2]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
  ),
  site_plots[[3]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
  ),
  nrow = 1,
  align = "h",
  labels = c(
    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
  )
)

mfitpls <- plot_grid(
  ar_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
  sitelumaps,
  ncol = 1,
  align = "v",
  rel_heights = c(1, .5),
  labels = c(
    "A", ""
  )
)

mfitpls_top <- plot_grid(
  sitelumaps + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
  ar_fit + theme(plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")),
  ncol = 1,
  align = "v",
  rel_heights = c(.5, 1),
  labels = c(
    "", "A"
  )
)

# all together
comb <- plot_grid(
  mfitpls_top, loading_plot,
  nrow = 1, align = "b",
  rel_widths = c(1, .4),
  labels = c("", "B", "")
)

comb

```

@@

# load land-use conversion table
luconversions <- read.csv("data/LUCategoryConversion.csv")
urban_luconversions <- luconversions %>%
  filter(LU == "urban")

@@

# make the maps for each of the sites selected in the previous step
site_shape_data_u <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/urban/",
    u_marked[i, "site"], "/", u_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  # add major LU categories
  df_merged <- df %>%
    mutate(descriptio = toupper(descriptio)) %>%
    full_join(urban_luconversions, by = c("descriptio" = "code"))
  site_shape_data_u[[u_marked[i, "site"]]] <- df_merged
  site_plots[[u_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df_merged,
      aes(fill = category)
    ) +
    theme_void() +
    scale_fill_manual(
      values = c(
        "continuous central" = "#EE6677", "dense residential" = "#AA3377",
        "sparse residential" = "#CCBB44", "amenity grassland" = "#228833",
        "water" = "#4477AA", "parks allotments cemeteries" = "#BBBBBB",
        "Railway" = "#BBBBBB", "woodland" = "#BBBBBB"
      )
    )
}

@@

loading_plot <- ggplot(urban.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(3)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.25,
    nudge_y = c(-0, -0), nudge_x = c(0.5, 0.5), size = rel(3)
  ) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None"
  ) +
  scale_colour_manual(
    values = c(
      "continuous.central" = "#EE6677", "dense.residential" = "#AA3377",
      "sparse.residential" = "#CCBB44", "amenity.grassland" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  ) +
  scale_y_continuous(breaks = seq(-.9, .9, by = 0.4), limits = c(-1, 1))

@@

# MAPS BELLOW PLOT
# first the maps in a column of three
sitelumaps <- plot_grid(
  site_plots[[1]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
  ),
  site_plots[[2]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
  ),
  site_plots[[3]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
  ),
  nrow = 1,
  align = "h",
  labels = c(
    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
  )
)

mfitpls <- plot_grid(
  u_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
  sitelumaps,
  ncol = 1,
  align = "v",
  rel_heights = c(1, .4),
  labels = c(
    "A", ""
  )
)

# all together
comb <- plot_grid(
  mfitpls, loading_plot,
  nrow = 1, align = "h",
  rel_widths = c(1, .8),
  labels = c("", "B", "")
)

comb

```

```{r}

vals <- c()
for (l in c(site_shape_data_ar, site_shape_data_u)) {
   vals <- c(vals, unique(as.vector(l$descriptio)))
}

unique(vals)

```
